---
title: "Examining Global Refugee and Internally Displaced Population Trends and Correlations (2010-2022)"
subtitle: "Proposal"
format: html
editor: visual
---

```{r}
#| label: load-pkgs
#| message: false
library(tidyverse)
```

## Dataset

```{r}
#| label: load-dataset
#| message: false
```

**Data-set Description:**\
The data-set comes from the {refugees} R package, which compiles information from important sources, including UNHCR, UNRWA, and IDMC, covering more than 70 years and a wide range of displacement circumstances. Data from 2010 to 2022 are included in this particular subgroup, which covers a wide range of characteristics including the country of origin, asylum names and codes, and the numbers of refugees, asylum seekers, returned refugees, internally displaced people, stateless people, and other individuals of concern.

**Reasons for Choosing this Data-set:**\
The significance of this data-set to current worldwide problems and displacement patterns—such as the effects of war, climate change, and political unrest on populations—was taken into consideration throughout its selection. This data-set was selected for another reason: its combination of categorical (such as country names, codes) and numerical (such as refugee and IDP counts) characteristics make it ideal for a wide variety of analytical and graphical techniques. Lastly, we're interested in learning the underlying motivations behind this, such as the effects of COVID, war, climate change, etc.

Make sure to load the data and use inline code for some of this information.

## Questions

Question 01:\
How have the trends in refugee populations changed over the years ? How the political parties view of USA effected this change? This question will utilize the 'year', 'coo_name', and 'refugees' variables.

## Analysis plan

-   A plan for answering each of the questions including the variables involved, variables to be created (if any), external data to be merged in (if any).
